Apparent preferences for cognitive effort fade when multiple forms of effort

### 2020-12-04 Cost_manuscript.Rmd
anova + plots for parameter comparison. now add paired permutations

``` {r OLD BTW and WTH: 16.3. OC comparisons btw, echo = FALSE, fig.align = "center", fig.width = 6, fig.height = 4}
# # ANOVA
# OCanova <- list()
# OCanova$aov <- summary(aov(Gamma ~ Cost, data = summaryOC_btw$all))
# Fval <- round(OCanova$aov[[1]]$`F value`[1], digits = 2) # doing separate variables because I can't easily use the `` in markdown.
# Pval <- OCanova$aov[[1]]$`Pr(>F)`[1]
# Rsquared_aov <- round(OCanova$aov[[1]]$`Sum Sq`[1] / sum(OCanova$aov[[1]]$`Sum Sq`), digits = 2)
#   
# # Post-hoc pairwise tests
# OCanova$perms <- summaryOC_btw$all %>%
#   select(Cost, Gamma) %>%
#   plyr::dlply("Cost", identity) %>%
#   do.call(cbind, .) %>%
#   select_if(is.numeric) %>%
#   rename(Physical = Physical.Gamma,
#          Cognitive = Cognitive.Gamma,
#          Wait = Wait.Gamma,
#          Easy = Easy.Gamma) %>%
#   outer(., ., FUN = Vectorize(permute), simple = T)
# diag(OCanova$perms) <- 1
# 
# OCanova$ES <- summaryOC_btw$all %>%
#   select(Cost, Gamma) %>%
#   plyr::dlply("Cost", identity) %>%
#   do.call(cbind, .) %>%
#   select_if(is.numeric) %>%
#   rename(Physical = Physical.Gamma,
#          Cognitive = Cognitive.Gamma,
#          Wait = Wait.Gamma,
#          Easy = Easy.Gamma) %>%
#   outer(., ., FUN = Vectorize(cohenD)) %>%
#   abs()
```

@@

``` {r BTW and WTH: 16.3. Model fits, echo = FALSE, fig.align = "center", fig.width = 6, fig.height = 4}
# the fitting of these models was performed by a separate script (modeling_gk_btw/wth)
# here we load the results to economize computational power during knitting
# load results from models
#load("models_gk/results/btw.Rdata")
#load("models_gk/results/wth.Rdata")
load("models_gk/results/mod_results.Rdata")

# keep relevant model results and combine into a single df with generalizable names (delete eventually)
temp_btw <- fits_bic_btw[c(1, 3, 4, 6), ]
temp_wth <- fits_bic_wth[c(1, 3, 13, 15), ]
temp_btw$wth <- temp_wth$bic
colnames(temp_btw) <- c("Model", "Between", "Within")
modelComparison <- temp_btw %>% gather(Exp, BIC, -Model)
rm(temp_btw, temp_wth, fits_bic_btw, fits_bic_wth)

@@

@@

  We tested four models against a set of criteria derived from observed choice behavior (see Methods for details). The first one, a base formulation, probed the predictions of the Marginal Value Theorem (Charnov, 1975). The second one allowed the opportunity cost to gradually adapt using a reinforcement learning rule. The third model assumed that different demands imposed distinct cost biases on the perceived opportunity costs, and that these biases could be recalibrated with experience. A final model combined the adaptive features of the second and third formulations.

>>>>

``` {r OLD BTW and WTH: 16.3. OC comparisons btw, echo = FALSE, fig.align = "center", fig.width = 6, fig.height = 4}
# ANOVA
OCanova <- list()
OCanova$aov <- summary(aov(Gamma ~ Cost, data = summaryOC_btw$all))
Fval <- round(OCanova$aov[[1]]$`F value`[1], digits = 2) # doing separate variables because I can't easily use the `` in markdown.
Pval <- OCanova$aov[[1]]$`Pr(>F)`[1]
Rsquared_aov <- round(OCanova$aov[[1]]$`Sum Sq`[1] / sum(OCanova$aov[[1]]$`Sum Sq`), digits = 2)

# Post-hoc pairwise tests
OCanova$perms <- summaryOC_btw$all %>%
  select(Cost, Gamma) %>%
  plyr::dlply("Cost", identity) %>%
  do.call(cbind, .) %>%
  select_if(is.numeric) %>%
  rename(Physical = Physical.Gamma,
         Cognitive = Cognitive.Gamma,
         Wait = Wait.Gamma,
         Easy = Easy.Gamma) %>%
  outer(., ., FUN = Vectorize(permute), simple = T)
diag(OCanova$perms) <- 1

OCanova$ES <- summaryOC_btw$all %>%
  select(Cost, Gamma) %>%
  plyr::dlply("Cost", identity) %>%
  do.call(cbind, .) %>%
  select_if(is.numeric) %>%
  rename(Physical = Physical.Gamma,
         Cognitive = Cognitive.Gamma,
         Wait = Wait.Gamma,
         Easy = Easy.Gamma) %>%
  outer(., ., FUN = Vectorize(cohenD)) %>%
  abs()
```

@@

``` {r BTW and WTH: 16.3. Model fits, echo = FALSE, fig.align = "center", fig.width = 6, fig.height = 4}
## the fitting of these models was performed by a separate script (modeling_gk_btw/wth)
# here we load the results to economize computational power during knitting
# load results from models
#load("models_gk/results/btw.Rdata")
#load("models_gk/results/wth.Rdata")
load("models_gk/results/mod_results.Rdata")

## BIC
# keep relevant model results and combine into a single df with generalizable names (delete eventually)
temp_btw <- fits_bic_btw[c(1, 3, 4, 6), ]
temp_wth <- fits_bic_wth[c(1, 3, 13, 15), ]
temp_btw$wth <- temp_wth$bic
colnames(temp_btw) <- c("Model", "Between", "Within")
modelComparison <- temp_btw %>% gather(Exp, BIC, -Model)
rm(temp_btw, temp_wth, fits_bic_btw, fits_bic_wth)

@@

## Model fit summaries
# mean and sd of parameters from winning model (exp some? still confused)
# parameter order for wth: gamma, alpha, C, w_decay, beta
mFits_btw <- sapply(fits_btw$`Adaptive OC + Biased OC`, function(group) {apply(group$x, 1, mean)})
sdFits_btw <- sapply(fits_btw$`Adaptive OC + Biased OC`, function(group) {apply(group$x, 1, sd)})

mFits_wth <- fits_wth$`Adaptive OC + Adaptive Between Biased OC`$mus
sdFits_wth <- diag(fits_wth$`Adaptive OC + Adaptive Between Biased OC`$sigma)

## Compare
# first, divide the group fits into a list of dataframes for each parameter
t <- lapply(fits_btw$`Adaptive OC + Biased OC`, function(group) {melt(group$x, value.name = "Fit")})
n <- names(fits_btw$`Adaptive OC + Biased OC`)
reshaped <- mapply(cbind, t, "Group" = n, SIMPLIFY = F)
aovShape_btw <- do.call(rbind, reshaped) %>% 
                  filter(Group != "pheasy") %>%
                  plyr::dlply("Var1", identity)
rm(t, n)

# then apply anovas
aovBtw <- lapply(aovShape_btw, function(d) {summary(aov(Fit ~ Group, data = d))})

## plot
# BTW
# select a parameter to plot
param <- aovShape_btw$cost
paramPlot <- ggplot(param, aes(Group, Fit, fill = Group)) +
  # geom_hline(yintercept = 0.64, alpha = 0.9, color = "gray20", size = 1) +
  # geom_hline(yintercept = 0.7, alpha = 0.9, color = "gray40", size = 1) +
  # geom_hline(yintercept = 0.9, alpha = 0.9, color = "gray60", size = 1) +
  geom_hline(yintercept = 0, alpha = 0.9, size = 1, linetype = "dashed") + # mean of highest earning rates across blocks
  #geom_dotplot(stackdir = 'center', binaxis = 'y', show.legend = F) +
  geom_boxplot(show.legend = F) +
  geom_jitter(width = 0.1, alpha = 0.5, show.legend = F, pch = 21, size = 3) +
  #ylim(0,1.5) +
  labs(x = "") +
  scale_fill_manual(values = colsBtw) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        text = element_text(size = 16))

@@

  We tested four models against a set of criteria derived from observed choice behavior (see Methods for details). The first one, a base formulation, probed the predictions of the Marginal Value Theorem (Charnov, 1975). The second one allowed individuals to estimate the opportunity cost at different rates using a reinforcement learning rule. The third model assumed that different demands imposed distinct cost biases on the experienced opportunity costs, and that these biases could be recalibrated with experience. A final model combined the adaptive features of the second and third formulations.

  The fourth model produced the lowest BIC in both experiments (Figure 10A. Between-subjects = $modelComparison[4, 3]$; Within-subjects = $modelComparison[8, 3]$).

  The fits of the model provided good qulitative resemblance to the observed choice behavior in both experiments (*Figure 10X*)
